---
title: "Statistical Inference Course Project. Part 2: Basic Inferential Data Analysis Instructionsless"
output: pdf_document
author: Vitalii Zhukov
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Description
Analyze the ToothGrowth data in the R datasets package.

## 1. Exploratory data analyses
ToothGrowth data loading:
```{r simulation}
library(datasets)
data("ToothGrowth")
```
Basic exploratory data analyses:
```{r }
head(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)
unique(ToothGrowth$len)
unique(ToothGrowth$supp)
unique(ToothGrowth$dose)
table(ToothGrowth$dose, ToothGrowth$supp) # split of cases between different (dose levels and delivery methods)

```


```{r plot1}
library(ggplot2)
##ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose))
ToothGrowth$dose = as.factor(ToothGrowth$dose)
g <- ggplot(ToothGrowth, aes(x=dose, y= len))
g + geom_boxplot(aes(fill=dose)) +facet_grid(.~supp) + ggtitle("Tooth Length vs. Doses, by suppl method")
```
Function of supplement type
```{r plot2}
g <- ggplot(ToothGrowth, aes(x=supp, y= len))
g + geom_boxplot(aes(fill=supp)) + facet_grid(.~dose) +
     ggtitle("Tooth Length vs. supplement method, by doses")
```

## Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose.
# t-test

Check for group differences due to different supplement type, assuming unequal variances between the two groups.
```{r }
t.test(len ~ supp, data = ToothGrowth)
```

The p-value is 0.06, and the confidence interval contains zero. This indicates that we can not reject the null hypothesis that the different supplement types have no effect on tooth length.

Create three sub-groups as per dose level pairs
```{r }
ToothGrowth.doses_0.5_1.0 <- subset (ToothGrowth, dose %in% c(0.5, 1.0)) 
ToothGrowth.doses_0.5_2.0 <- subset (ToothGrowth, dose %in% c(0.5, 2.0)) 
ToothGrowth.doses_1.0_2.0 <- subset (ToothGrowth, dose %in% c(1.0, 2.0)) 
```

Check for group differences due to different dose levels (0.5, 2.0), assuming unequal variances between the two groups
```{r }
t.test(len ~ dose, data = ToothGrowth.doses_0.5_2.0) 
```

Check for group differences due to different dose levels (0.5, 1.0), assuming unequal variances between the two groups
```{r }
t.test(len ~ dose, data = ToothGrowth.doses_0.5_1.0)
```

Check for group differences due to different dose levels (1.0, 2.0),assuming unequal variances between the two groups
```{r }
t.test(len ~ dose, data = ToothGrowth.doses_1.0_2.0)
```

For all three dose level pairs, the p-value is less than 0.05, and the confidence interval does not contain zero. The mean tooth length increases on raising the dose level. This indicates that we can reject the null hypothesis, and establish that increasing the dose level leads to an increase in tooth length.

## Conclusions
Supplement type has no effect on tooth growth.
Inreasing the dose level leads to increased tooth growth.

## Assumptions
The experiment was done with random assignment of guinea pigs to different dose level categories and supplement type to control for confounders that might affect the outcome.
Members of the sample population, i.e. the 60 guinea pigs, are representative of the entire population of guinea pigs. This assumption allows us to generalize the results.
For the t-tests, the variances are assumed to be different for the two groups being compared. This assumption is less stronger than the case in which the variances are assumed to be equal.